---
title: "Halloween Action Data"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    df_print: paged
---

```{r}
library(tidyverse)
```

```{r}
candy <- read_csv("clean_data/candy.csv")

```
## Cleaning and assumptions

It was assumed that 'candy' was being used to refer to anything
received while trick or treating.
It was also assumed that anything not in the above category or
specified later was to be unnecessary, and could be dropped.
For any responses which did not take the expected form, for example
text responses when asked their age, it was assumed that the respondent
didn't wish to answer the question.
As a continuation of this, any countries given which were not obviously
alternate names or misspellings are left as given.

## 1 What is the total number of candy ratings given across the three years?
```{r}
nrow(candy)
```

## 2 What was the average age of people who are going out trick or treating?
```{r}
candy %>% 
  filter(going == TRUE) %>% 
  distinct(id, .keep_all = TRUE) %>% 
  summarise(avg_age = mean(age, na.rm = TRUE))
```

## 3 What was the average age of people who are not going trick or treating?
```{r}
candy %>% 
  filter(going == FALSE) %>% 
  distinct(id, .keep_all = TRUE) %>% 
  summarise(avg_age = mean(age, na.rm = TRUE))
```

## 4 For each of joy, despair and meh, which candy bar received the most of these ratings?
```{r}
candy %>% 
  group_by(question) %>% 
  summarise(
    despair = sum(response == "DESPAIR"),
    joy = sum(response == "JOY"),
    meh = sum(response == "MEH")) %>% 
  mutate(
    max_despair = max(despair),
    max_joy = max(joy),
    max_meh = max(meh)
    ) %>% 
  filter(
    max_despair == despair |
      max_joy == joy |
      max_meh == meh
  ) %>% 
  select(question, joy, despair, meh)
```

## 5 How many people rated Starburst as despair?
```{r}
candy %>% 
  filter(question == "starburst") %>% 
  summarise(starburst_haters = sum(response == "DESPAIR"))
```
## rating system
count despair as -1, joy as +1, and meh as 0.

## 6 What was the most popular candy bar by this rating system for each gender in the dataset ?
```{r}
candy %>% 
  group_by(gender, question) %>% 
  summarise(popularity = sum(response == "JOY") - sum(response == "DESPAIR")) %>% 
  mutate(most_popular = max(popularity)) %>% 
  filter(popularity == most_popular) %>% 
  select(gender, question, popularity)
```

## 7 What was the most popular candy bar in each year?
```{r}
candy %>% 
  group_by(year, question)%>% 
  summarise(popularity = sum(response == "JOY") - sum(response == "DESPAIR")) %>% 
  mutate(most_popular = max(popularity)) %>% 
  filter(popularity == most_popular) %>% 
  select(year, question, popularity)
```

## 8 What was the most popular candy bar by this rating for people in US, Canada, UK, and all other countries?
Here, an assumption was made that countries with fewer than 6
respondents were not represented in the data, and were not included.

```{r}
candy_country <- candy %>% 
  distinct(id, .keep_all = TRUE) %>% 
  group_by(country) %>% 
  summarise(individuals = n()) %>% 
  filter(individuals > 5) %>% 
  pull(country)

candy %>% 
  filter(country %in% candy_country) %>% 
  group_by(country, question) %>% 
  summarise(popularity = sum(response == "JOY") - sum(response == "DESPAIR")) %>% 
  mutate(most_popular = max(popularity)) %>% 
  filter(popularity == most_popular) %>% 
  select(country, question, popularity)
```

